Ages of the Kepler Stars and Planets

### 2024-04-25 paper/ms.tex
update ms

@@

{\it RUWE (bit 7)}---We flagged stars with a Gaia DR3 renormalized
unit weight error
(RUWE\footnote{\url{https://gea.esac.esa.int/archive/documentation/GDR3/Gaia_archive/chap_datamodel/sec_dm_main_source_catalogue/ssec_dm_gaia_source.html}})
exceeding 1.4 as possible binaries.  Such astrometric outliers can be
either bona fide astrometric binaries, or more often are marginally
resolved point sources for which the single-source PSF model used in
the Gaia pipeline induces spurious astrometric scatter.

@@

\begin{figure}[!t]
	\begin{center}
		\subfloat{
			\includegraphics[width=0.48\textwidth]{prot_teff_Santos19_Santos21_dquality.pdf}
		}

		\vspace{-0.5cm}
		\subfloat{
			\includegraphics[width=0.48\textwidth]{koi_mean_prot_teff_koi_X_S19S21dquality_keepgrazing.pdf}
		}
	\end{center}
	\vspace{-0.5cm}
	\caption{
    {\bf Rotation periods for Kepler stars (top) and planets
    (bottom)}.  The periods are from \citet{Santos_2019} and
    \citet{Santos_2021} (\citetalias{Santos_2019} and
    \citetalias{Santos_2021}).  The gray lines are ``mean fits'' from
    Figure~\ref{fig:models} to the slow rotation sequences of M67,
    Ruprecht~147, NGC~6811, Praesepe, and a trio of 120\,Myr open
    clusters.  The stellar sample shows only apparently single stars
    near the main sequence with $\log g$$>$4.2, RUWE$<$1.4,
    3800$<$$T_{\rm eff}$/K$<$6200, that are not flagged as eclipsing
    binaries.  The planet sample has the same stellar cuts, and shows
    confirmed and candidate planets from the cumulative KOI table,
    requiring them to have a MES$>$10.
	}
	\label{fig:prot_vs_teff}
\end{figure}

@@

\startlongtable
\begin{deluxetable*}{lllllllrrrrrr}
	%\tabletypesize{\footnotesize}
	\tabletypesize{\scriptsize}
  \tablecaption{Ages of Kepler target stars.  
  The full machine-readable table contains rotation-based ages
  for \nuniqstarsantosrotteffcut\ stars with temperatures spanning
  3800-6200\,K.
  \nuniqstarsantosrotgyroappl\ of these stars have quality flags
  without bits 0 to 9 raised, meaning that they are likely valid
  objects for gyrochronology (see Section~\ref{subsec:flags}).
  \label{tab:stars}}
	%\toprule
	%\midrule
	%\endhead
	\startdata
  KIC & Gaia &  $T_{\rm eff}$ & $P_{\rm rot}$ & $t_{\rm gyro}$ & $Q_{\rm gyro}$ \\
  -- &   -- & K & days &  Myr &    int  \\
	\hline
  %FIXME make it!
	%\input{table_gyrostars} %\\
	%\hline
	\enddata
%	\tablecomments{This table includes X, Y, Z... The machine-readable version,
%		available online, includes additional columns for .... 
%	}
\end{deluxetable*}

>>>>

\begin{figure*}[!t]
	\begin{center}
		\subfloat{
			\includegraphics[width=0.48\textwidth]{prot_teff_Santos19_Santos21_dquality.pdf}
		}
		%	
		%		\vspace{-0.5cm}
		\subfloat{
			\includegraphics[width=0.48\textwidth]{koi_mean_prot_teff_koi_X_S19S21dquality_keepgrazing.pdf}
		}
	\end{center}
	\vspace{-0.5cm}
	\caption{
		{\bf Rotation periods for Kepler stars (top) and planets
			(bottom)}.  The periods are from \citet{Santos_2019} and
		\citet{Santos_2021} (\citetalias{Santos_2019} and
		\citetalias{Santos_2021}).  The gray lines are ``mean fits'' from
		Figure~\ref{fig:models} to the slow rotation sequences of M67,
		Ruprecht~147, NGC~6811, Praesepe, and a trio of 120\,Myr open
		clusters.  The stellar sample shows only apparently single stars
		near the main sequence with $\log g$$>$4.2, RUWE$<$1.4,
		3800$<$$T_{\rm eff}$/K$<$6200, that are not flagged as eclipsing
		binaries.  The planet sample has the same stellar cuts, and shows
		confirmed and candidate planets from the cumulative KOI table,
		requiring them to have a MES$>$10.
	}
	\label{fig:prot_vs_teff}
\end{figure*}

@@

{\it RUWE (bit 7)}---We flagged stars with a Gaia DR3 renormalized
unit weight error
(RUWE) %\footnote{\url{https://gea.esac.esa.int/archive/documentation/GDR3/Gaia_archive/chap_datamodel/sec_dm_main_source_catalogue/ssec_dm_gaia_source.html}})
exceeding 1.4 as possible binaries.  Such astrometric outliers can be
either bona fide astrometric binaries, or more often are marginally
resolved point sources for which the single-source PSF model used in
the Gaia pipeline induces spurious astrometric scatter.

@@

@@

\startlongtable
\begin{deluxetable*}{lllllllrrrrrr}
	\tabletypesize{\scriptsize}
  \tablecaption{
  Ages of Kepler target stars derived from rotation periods.  The
  full machine-readable table includes \nuniqstarsantosrot\ stars
  with reported rotation periods from \citetalias{Santos_2019} and
  \citetalias{Santos_2021}, \nuniqstarfinitegyroage\ of which have
  finite reported gyrochrone ages.  A random set of these stars is
  shown here to give an idea for form and content.  The quality
  bit-flag is as in Table~\ref{tab:planets}; requiring bits zero to
  nine to be null yields \nuniqstarsantosrotgyroappl\ stars for
  which gyrochronology is likely to be valid.  \label{tab:stars}}
	%\toprule
	%\midrule
	%\endhead
	\startdata
  KIC & Gaia DR3 &  $T_{\rm eff}$ & $P_{\rm rot}$ & $t_{\rm gyro}$ & $Q_{\rm gyro}$ \\
  -- &   -- & K & days &  Myr &    int  \\
	\hline
	\input{table_star_gyro} %\\
	%\hline
	\enddata
%	\tablecomments{This table includes X, Y, Z... The machine-readable version,
%		available online, includes additional columns for .... 
%	}
\end{deluxetable*}
